\sysfull

### 2018-08-06 paper/asplos19/paging.tex

%\note{additional motivation for coala memory virtualization] we should emphasize that Coala programming model removes the burden of identifying the write-after-read dependencies from the programmer  as compared to Chain. And it does not require a specific compiler that cannot consider a runtime dependent behavior like Interrupts service handling to guarantee memory protection, as in Alpaca.}
%
\sys's {\em virtual memory manager} (VMM) is a core enabler of \sys task adaptation mechanism.
Its overview is given in Figure~\ref{figure:mem-man}.
VMM abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers, and each buffer is divided into pages.
\texttt{\underline{private}} holds the consistent version of pages after each commit (and on reboot), while \texttt{\underline{shadow}} serves as support during the execution of coalesced tasks.
The VMM prohibits applications from directly accessing these buffers.
Instead, it redirects any request to the \texttt{working} buffer located in the volatile memory (SRAM), which has a lower latency and energy cost to access than FRAM.
It populates the \texttt{working} buffer with the privatized pages from non-volatile memory requested by tasks.
When a coalesced task completes, the VMM ensures that {\em dirty} (i.e., modified) pages are committed to FRAM and visible to subsequent tasks.
If a power failure interrupts a task, all temporary modifications to pages in \texttt{working} (and \texttt{\underline{shadow}}) occurred after the last commit are discarded in order to keep data consistency.
% If the power interrupts a task, the VMM ensures the consistency of the data by relying on its \texttt{\underline{private}} buffer. The VMM ensures consistency even if a task or a sequence of tasks include WAR dependences that could compromise consistency~\cite{ratchet,dino}.
On a commit, the VMM atomically copies, through a two-phase commit, all dirty pages back into their location in the non-volatile memory.
% SRAM has a limited size and there is an extended working memory containing {\em shadow pages} in the larger FRAM. A shadow page stores the data from an updated page that needs to be committed back to non-volatile memory, but that exceeded the size of the volatile working memory.

>>>>

%\note{additional motivation for coala memory virtualization] we should emphasize that Coala programming model removes the burden of identifying the write-after-read dependencies from the programmer  as compared to Chain. And it does not require a specific compiler that cannot consider a runtime dependent behavior like Interrupts service handling to guarantee memory protection, as in Alpaca.}

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/coalesced-war.pdf}
	\caption{induction of WAR dependency by coalescing: without extra protection task coalescing induces WAR dependency (at variable $x$) even though individual tasks are WAR protected.}
	\label{figure:coalescedWar}
\end{figure}

\sys's {\em virtual memory manager} (VMM) is a core enabler of \sys task adaptation mechanism. Its necessity is illustrated by example given in Figure~\ref{figure:coalescedWar}: even though each task might be WAR-protected, coalescing breaks this insurance. Therefore we need a new mechanism that will keep WAR-protection despite coalescing.

VMM overview is given in Figure~\ref{figure:mem-man}.
VMM abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers, and each buffer is divided into pages.
\texttt{\underline{private}} holds the consistent version of pages after each commit (and on reboot), while \texttt{\underline{shadow}} serves as support during the execution of coalesced tasks.
The VMM prohibits applications from directly accessing these buffers.
Instead, it redirects any request to the \texttt{working} buffer located in the volatile memory (SRAM), which has a lower latency and energy cost to access than FRAM.
It populates the \texttt{working} buffer with the privatized pages from non-volatile memory requested by tasks.
When a coalesced task completes, the VMM ensures that {\em dirty} (i.e., modified) pages are committed to FRAM and visible to subsequent tasks.
If a power failure interrupts a task, all temporary modifications to pages in \texttt{working} (and \texttt{\underline{shadow}}) occurred after the last commit are discarded in order to keep data consistency.
% If the power interrupts a task, the VMM ensures the consistency of the data by relying on its \texttt{\underline{private}} buffer. The VMM ensures consistency even if a task or a sequence of tasks include WAR dependences that could compromise consistency~\cite{ratchet,dino}.
On a commit, the VMM atomically copies, through a two-phase commit, all dirty pages back into their location in the non-volatile memory.
% SRAM has a limited size and there is an extended working memory containing {\em shadow pages} in the larger FRAM. A shadow page stores the data from an updated page that needs to be committed back to non-volatile memory, but that exceeded the size of the volatile working memory.
